---
title: " Data Science Capstone project"
author: "Nikhitha Karlapudi"
date: '`r Sys.Date()`'
format:
  html:
    code-fold: true
course: STA 6257 - Advanced Statistical Modeling!
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

## Introduction

K-means clustering   is one of the most commonly used unsupervised machine learning algorithm for partitioning a given data set into a set of k groups (i.e. k clusters), where k represents the number of groups pre-specified by the analyst. It classifies objects in multiple groups (i.e., clusters), such that objects within the same cluster are as similar as possible (i.e., high intra-class similarity), whereas objects from different clusters are as dissimilar as possible (i.e., low inter-class similarity). In k-means clustering, each cluster is represented by its center (i.e, centroid) which corresponds to the mean of points assigned to the cluster.

### What is "method"?

K-means algorithm requires users to specify the number of cluster to generate.

The R functions are used to estimate the clusters and finally we have to visualize the K-means clusters.

One of the advantage is K-means clustering is very simple and fast and disadvantage is it requires some prior knowledge for selecting the clusters.

Kmeans clustering involves the following steps

1.selection of clusters 

2.Randomly picking the centroids

3.Thirdly, assigning observations to centroid it belongs based on similarity

4.Calculating the centroid and repeating the steps until the maximum number of iterations are reached 

### Related work

This section is going to cover the literature review...

## Methods

The common non-parametric regression model is
$Y_i = m(X_i) + \varepsilon_i$, where $Y_i$ can be defined as the sum of
the regression function value $m(x)$ for $X_i$. Here $m(x)$ is unknown
and $\varepsilon_i$ some errors. With the help of this definition, we
can create the estimation for local averaging i.e. $m(x)$ can be
estimated with the product of $Y_i$ average and $X_i$ is near to $x$. In
other words, this means that we are discovering the line through the
data points with the help of surrounding data points. The estimation
formula is printed below [@R-base]:

$$
M_n(x) = \sum_{i=1}^{n} W_n (X_i) Y_i  \tag{1}
$$ $W_n(x)$ is the sum of weights that belongs to all real numbers.
Weights are positive numbers and small if $X_i$ is far from $x$.

Another equation:

$$
y_i = \beta_0 + \beta_1 X_1 +\varepsilon_i
$$

## Analysis and Results

### Data and Visualization

A study was conducted to determine how...

```{r, warning=FALSE, echo=T, message=FALSE}
# loading packages 
library(tidyverse)
library(knitr)
library(ggthemes)
library(ggrepel)
library(dslabs)
```

```{r, warning=FALSE, echo=TRUE}
# Load Data
kable(head(murders))

ggplot1 = murders %>% ggplot(mapping = aes(x=population/10^6, y=total)) 

  ggplot1 + geom_point(aes(col=region), size = 4) +
  geom_text_repel(aes(label=abb)) +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(formula = "y~x", method=lm,se = F)+
  xlab("Populations in millions (log10 scale)") + 
  ylab("Total number of murders (log10 scale)") +
  ggtitle("US Gun Murders in 2010") +
  scale_color_discrete(name = "Region")+
      theme_bw()
  

```

### Statistical Modeling

```{r}

```

### Conclusion
K-means is one method of cluster analysis that groups observations by minimizing Euclidean distances between them. Euclidean distances are analagous to measuring the hypotenuse of a triangle, where the differences between two observations on two variables (x and y) are plugged into the Pythagorean equation to solve for the shortest distance between the two points

## References
https://www.datanovia.com/en/lessons/k-means-clustering-in-r-algorith-and-practical-examples/
https://www.datacamp.com/tutorial/k-means-clustering-r
[@Zubair M, Iqbal MDA, Shil A, Chowdhury MJM, Moni MA, Sarker IH. An Improved K-means Clustering Algorithm Towards an Efficient Data-Driven Modeling. Ann. Data. Sci. 2022 Jun 25:1–20. doi: 10.1007/s40745-022-00428-2. Epub ahead of print. PMCID: PMC9243813.
 simplilearn.com]
https://www.publichealth.columbia.edu/research/population-health-methods/k-means-cluster-analysis
